---
title: "lab_02"
author: "derek willis"
date: "2023-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1** Create a new column that divides the expense amounts into the following categories - \$0-\$500, \$500-\$5000, \$5001-\$50000, \$50001\_\$100000, \$100001-\$500000 - giving each an appropriate text value, and count how many expenses are in each group. How would you describe the distribution of expenses in a sentence? Would you change the categories?

**A1.** The majority of the expenses are level 3 expenses and fall between $5001 to $50000. I would divide the expenses into even smaller intervals, since there's a large range difference between just $0 to $500 and $5001 to $50000.

```{r}
maryland_expenses %>%
  mutate(
    expense_level = case_when(
     amount >=0 & amount <=500 ~ "Level 1 Expense",
     amount >=501 & amount <=5000 ~ "Level 2 Expense",
     amount >=5001 & amount <=50000 ~ "Level 3 Expense",
     amount >=50001 & amount <=100000 ~ "Level 4 Expense",
     amount >=100001 & amount <=500000 ~ "Level 5 Expense",
    )
  ) %>%
  group_by(expense_level) %>%
  summarise(
    total_amount = sum(amount),
  )

```

**Q2.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? 
**A2.** 26 departments, Atmospheric and Oceanic Science offers the most at 65 classes

```{r}
umd_courses %>% 
  filter(str_detect(title, "Climate"), seats>=1) %>% 
  group_by(department) %>% 
  summarise(
    count_classes = n()
  ) %>%
  arrange(desc(count_classes))
```

**Q3.** Of the departments listed in A2, which one is the most interesting to you, and why?
**A3.** Criminology and Criminal Justice - as a crim major myself, we never bring up climate whether that's the actual climate of the planet or a political/social climate, so I'm wondering if any of the classes included in this count actually have "climate" in the title.

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.
**A4.** Wes Moore's committee seems to spend the most money in October, the month right before the gubernatiorial election took place. The committee also spent more money in June and July, which were the months leading up to the primary to select candidates.

```{r}
maryland_expenses <- maryland_expenses %>% mutate(expenditure_date=mdy(expenditure_date))
```
```{r}
maryland_expenses %>%
  mutate(month = floor_date(expenditure_date, "month")) %>%
  group_by(month) %>%
  summarise(
    total_amount = sum(amount),
    ) %>%
  arrange(desc(total_amount))
```
```{r}
moore_committee_donations <- maryland_expenses %>% filter(committee_name == "Moore  Wes For Maryland")
```
```{r}
moore_committee_donations %>%
  mutate(month = floor_date(expenditure_date, "month"))  %>%
 group_by(month) %>%
  summarise(
    total_amount = sum(amount),
    ) %>%
  arrange(desc(total_amount))
```


**Q5.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.
**A5.** Question: Which expense purpose was the most frequently cited?
The most frequently cited expense_purpose was Credit Card Service Processing Charges.

```{r}
maryland_expenses %>%
  group_by(expense_purpose) %>%
  summarise(
    count_expense_purpose = n(),
  ) %>%
arrange(desc(count_expense_purpose))
```
