---
title: "lab_12"
author: "derek willis"
date: "2023-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
library(rvest)
```


## Questions

**Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

**A1.** The leading word according to the results is failing.  The top 10 words seem like they could be describing legal issues dealing with clients, accounts, consent, and something failing, especially given that failing is the top word.

```{r}
md_attorney_sanctions <- read_csv('data/md_attorney_sanctions.csv')

unique_words <- md_attorney_sanctions %>% select(text) %>% unnest_tokens(word, text)

unique_words %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

**Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

**A2.** The top trigram is attorney trust account, which, in legal terms, is an account where attorneys keep client funds safe until they need to be withdrawn. The top trigram appears 343 times.

```{r}
md_attorney_sanctions %>%
  unnest_tokens(trigram, text, token = "ngrams", n = 3) %>%
  separate(trigram, c("word1", "word2", "word3"), sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  filter(!word3 %in% stop_words$word) %>%
  mutate(trigram = paste(word1, word2, word3, sep=" ")) %>%
  group_by(trigram) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

**Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

**A3.** The number of occurrences went up significantly beginning in 2017, and through 2019-2022, the number of occurrences of this trigram was between 20-30 times in the sanctions for these fiscal years. I'd believe this suggests that during these fiscal years, there were significantly more sanctions involving misconduct surrounding attorney trust accounts. I'd be interested in looking further into these fiscal years specifically and looking for trigrams during these years. If there's a high number of occurrences for some trigrams, that could be a hint as to what misconduct involving the attorney trust accounts actually was. To clarify other important reasons attorneys get sanctioned, I could conduct some online research for these reasons and then search for phrases relevant to those reasons in R to show if they appear in the sanctions.

```{r}
md_attorney_sanctions %>% 
  filter(str_detect(text, "attorney trust account")) %>% 
  group_by(fiscal_year) %>%
  summarise(
    count=n()
  )
```
