---
title: "Class Reference"
author: "Gabrielle Sophia Lewis"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example:

```{r}
setwd("~/Documents/GitHub/data_journalism_2023_spring")
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### Tidyverse Tools

Lubridate -> helps working with date-times and time-spans, can mutate columns into dates that R can interpret, can create new columns from data
  * functions: mdy, ymd, ymd_hms, floor_date (groups all dates into the first one mentioned of a specific month/year), str_squish() to
    clean up any extra space, str_replace_all(column name, 'thing to replace', 'replacement'), wday(df$date column, label = TRUE) to get      weekday labels from date column, bind_rows(list(dataframe 1, dataframe 2, etc.)) and bind_cols(dataframe 1, dataframe 2, etc.)) to 
    combine files with IDENTICAL structures of data, left_join(added dataframe, by ="column name")
    
Janitor -> helps clean up data, can standardize column names (ex. Fiscal Year -> fiscal_year)
  * functions: clean_names(), rename(new name = old name), get_dupes() to find duplicates, distinct() to get rid of ALL duplicates
  
Rvest -> webscrape from HTML addresses to get data
  * functions: read_html() to read html link, html_elements('element'), html_text(), html_table()

### Codeblock Commands/Functions

glimpse() -> get a sense of the data, some of the values in each field/variable/column, and the data types of each field/variable/column.
group_by() -> group the data by a specific condition
summarise() -> summarize the data with functions contained within the ()
guess_max() -> guesses what type of column is in the data based on the values in the column; value in the () is how many rows it'll use to guess
problems(variable) -> tells you the issues with a variable

n() -> counts rows of data
nrow(dataframe name) -> shows # in a certain dataframe
drop_na(column name) -> drops any rows with NA value

head() -> show first n rows of data

str_detect(column, "term") -> string detect, is this phrase in this column?
  * "^term" -> does this phrase show up at the START of the column?
  * "term^" -> does this phrase show up at the END of the column?
str_to_upper(column name) -> convert a character column into all uppercase, can standardize letter case
case_when() -> can handle multiple variations, says when the data shows a specific scenario to do a specific action
str_length() -> string length, is this length in this column?
  
mutate(column name=function(column name)) -> can change column type

slice(row #) -> cuts out a certain row from data frame

col_names = FALSE -> if R reads a row of data as column names, this tells them that the row is actually data

### Filters vs. Selections

filter(name == "value") is for ROWS; create a new variable to hold the subset you're picking out
select() is for COLUMNS

### Abbreviations/Meanings

%>% and then do this
obs means rows
variables means columns
== exact matching
!= does not match
AND is `&` and OR is `|`, format of whatever is on the sides of &/| needs to be identical
> or >=, < or <=

### Assigning values to variable names

Use a left-pointing arrow (assignment operator) to put a value to a variable. Run these in the console.

firstnumber <- 1
secondnumber <-2

### How to make comments

Put a # in front of a sentence in a codeblock to create a comment that won't run with the block.

```{r}
# examples
library(tidyverse)
```

