---
title: "Class Reference"
author: "Gabrielle Sophia Lewis"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example:

```{r}
setwd("~/Documents/GitHub/data_journalism_2023_spring")
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### Codeblock Commands/Functions
glimpse() = get a sense of the data, some of the values in each field/variable/column, and the data types of each field/variable/column.
group_by() = group the data by a specific condition
summarise() = summarize the data with functions contained within the ()


### Abbreviations

obs means rows
variables means columns

### Assigning values to variable names

Use a left-pointing arrow (assignment operator) to put a value to a variable. Run these in the console.

firstnumber <- 1
secondnumber <-2

### How to make comments

Put a # in front of a sentence in a codeblock to create a comment that won't run with the block.

```{r}
# examples
library(tidyverse)
```

