---
title: "Reverse Engineering Project"
author: "Gabrielle Lewis, Damon Brooks"
date: "2023-03-04"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Why Many Americans Don't Vote](https://projects.fivethirtyeight.com/non-voters-poll-2020-election/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
nonvoters_data <- read_csv("data/nonvoters_data.csv")

# Clean required data and prepare for analysis if needed. 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: People who sometimes vote were a plurality of the group (44 percent), while 31 percent nearly always cast a ballot and just 25 percent almost never vote."
* **Analysis summary**: The results reflect what was written in story, which is that out of all the survey respondents, about 44% — the majority — were sporadic voters, about 31% were frequent voters, and about 25% almost never voted.

```{r}
# Put code to reverse engineer sentence here
nonvoters_data %>%
  group_by(voter_category) %>%
  summarise(
    count=n()
  ) %>%
  mutate(
    category_percentage=(count/sum(count))*100
  ) %>%
  arrange(desc(count))
# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: "Nonvoters were more likely to have lower incomes; to be young; to have lower levels of education; and to say they don’t belong to either political party, which are all traits that square with what we know about people less likely to engage with the political system."

The sentence asks us to find the percentages of each voter category group that reported the lowest income bracket, the youngest age group, the lowest level of education, and not identifying with either major political party in the survey.

* **Analysis summary**: Based on my findings, I’ve learned that the voter_category titled “Rarely/never” with a college education has the lowest count of 423. While, the voter category titled “sporadic” with a college education has the highest count of 1073. Furthermore, my findings have shown me that having a college education will translate into a higher income, regardless of voting category. I think we were able to confirm the findings because we have sufficient to support the results. 

```{r}
# Put code to reverse engineer sentence here

# lower incomes
nonvoters_data %>%
  group_by(income_cat, voter_category) %>%
  summarise(
    count=n()
  ) %>%
    mutate(
    percent = count/sum(count) * 100
  )

# age
nonvoters_data %>%
  group_by(ppage, voter_category) %>%
  summarise(
    count=n()
  ) %>%
    mutate(
    percent = count/sum(count) * 100
  )

# education level
nonvoters_data %>%
  group_by(educ, voter_category) %>%
  summarise(
    count=n()
  ) %>%
    mutate(
    percent = count/sum(count) * 100
  )

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: "For instance, people who vote more irregularly are a little more likely than those who almost always vote to have lost a job in the past year (13 percent vs. 10 percent) or to have worried about expenses (21 percent vs. 16 percent), although those who almost never vote are even likelier than those who sometimes vote to have reported these experiences."

The sentence seems to be asking for code that shows the number of occurrences of each response (yes = 1 and no = 2) under each voter category, then converting that number into a percentage of the voter category group.

* **Analysis summary**: About 10% of people who always vote, about 13% of people who sometimes vote and about 18% of people who rarely vote worried about a lost job at the time of this story. This lines up with what the story said that those who rarely vote are more likely to report having lost a job than those who always or sometimes vote. For the next question, about 15% of people who always vote, about 21% of people who sometimes vote, and about 33% of people who rarely vote reported worrying about expenses. This lines up with the story's sentence that people who rarely vote report this concern more than the other two voting groups, but the percentage of people who always vote that we found is 15.4, which doesn't technically round up to the 16% written in the story. The 21% finding remains accurate.


```{r}
# Put code to reverse engineer sentence here
nonvoters_data %>%
  group_by(voter_category) %>%
  count(Q11_1) %>%
  mutate(
    pct_lostjob_always=(179/1811)*100,
    pct_lostjob_sporadic=(330/2574)*100,
    pct_lostjob_never=(258/1451)*100
  )

# Display results of code below this codeblock

```
```{r}
nonvoters_data %>%
  group_by(voter_category) %>%
  count(Q11_5) %>%
  mutate(
    pct_expenses_always=(279/1811)*100,
    pct_expenses_sporadic=(541/2574)*100,
    pct_expenses_never=(472/1451)*100
  )
```


### Sentence 4

* **Sentence text**: "Our survey found, for instance, that occasional voters were slightly more likely than frequent voters to have a long-term disability (8 percent vs. 5 percent), and nonvoters were even more likely to fall into this category (12 percent)."
* **Analysis summary**: We were able to confirm the findings. We learned about 8% of sporadic voters reported a long term disability. About 5% of voters who always vote reported that they have a long term disability. Roughly 12% of voters who say they rarely or never vote reported that they have a long term disability.

```{r}
# Put code to reverse engineer sentence here
nonvoters_data %>%
  group_by(voter_category, Q10_1) %>%
  summarise(responses=n()) %>%
  mutate(New_percent = responses/sum(responses)*100)

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: "Others agreed that there’s more the U.S. could be doing to encourage people to vote. Fifty-one percent of respondents thought that making Election Day a national holiday would allow more people to cast a ballot. Similar shares agreed that more information about the candidates from unbiased sources (51 percent) or early in-person voting (52 percent) would help, too."

The sentence is asking for us to use code to figure out the percentage of "Yes" responses, or a response of '1', for each of the questions.

* **Analysis summary**: The findings from our own code almost lined up with the story's findings. 51% of respondents said info from unbiased sources would be helpful getting people to vote. However, the other percentages we found were slightly higher. Instead of 51% percent responding to make Election Day a national holiday, about 52% responded, and instead of 52% responding that early in-person voting would help, about 53% — which could be even rounded up to 54% — responded.

```{r}
# Put code to reverse engineer sentence here
nonvoters_data %>%
  group_by(Q19_2) %>%
  summarise(
    responses = n()
    ) %>%
  mutate(Freq = responses/sum(responses)*100)

nonvoters_data %>%
  group_by(Q19_3) %>%
  summarise(
    responses = n()
    ) %>%
  mutate(Freq = responses/sum(responses)*100)

nonvoters_data %>%
  group_by(Q19_6) %>%
  summarise(
    responses = n()
    ) %>%
  mutate(Freq = responses/sum(responses)*100)
# Display results of code below this codeblock

```

-30-