---
title: "lab_03"
author: "Derek Willis"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2023.csv", guess_max=322138)

# Load in the Maryland grants and loan data
md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")

# Cleaning up names
cleaned_md_grants_loans <- md_grants_loans %>%
  clean_names()

cleaned_payments <- payments %>%
  clean_names()
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: I found a total of 79 records.

```{r}
pg_schools <- cleaned_payments %>%
  group_by(vendor_name) %>%
  filter(
    str_detect(vendor_name, "PRINCE GEORGE"), str_detect(vendor_name, "SCHOOL") | str_detect(vendor_name, "BOARD") 
    )
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed?

A2: 2022 had the largest total amount at $1,578,857,325. The fiscal years span from 2008 to 2023, but some years are missing such as 2011 and 2014.

```{r}
pg_schools %>%
  group_by(fiscal_year) %>%
  summarise(
    total_amount=sum(amount)
  ) %>%
  arrange(desc(total_amount))

```
Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3. To clean the data so nearly identical vendors are combined, my first thought is to mutate these vendors' names and use the case_when() function. Examples of vendors with name variations throughout the data are Plano-Coudon, LLC (variation of Plano-Coudon LLC), A Morton Thomas & Associates Inc (variation of A Morton Thomas & Assoc Inc) and Olympus America Inc. (variation of Olympus America Inc). For vendors like these, the case_when() function would show if a row contains key terms found in all the variations — for example, both Olympus America Inc. variations have "Olympus America Inc" no period — in that case, they would both be given the same name to make the identical.

```{r}
cleaned_payments %>%
  group_by(vendor_name) %>%
  filter(
    str_detect(agency_name, "UNIVERSITY OF MARYLAND")
    ) %>%
  summarise(
    total_amount=sum(amount)
  ) %>%
  arrange(vendor_name)
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4. Since 2009 — which has the highest amount of grants for stem cell research projects out of all the fiscal years in the data — the amount of grants given to these projects has steadily decreased. The grants hit their lowest point in this data in 2021 but then shot back up in 2022 to a similar amount as 2020.

```{r}
cleaned_md_grants_loans <- cleaned_md_grants_loans %>%
  mutate(
    grantor=str_to_upper(grantor), 
    grantee=str_to_upper(grantee),
    description=str_to_upper(description)
    )

cleaned_md_grants_loans %>%
  group_by(fiscal_year) %>%
  filter(
    str_detect(description, "STEM CELL")
    ) %>%
  summarise(
    total_amount=sum(amount),
  ) %>%
  arrange(fiscal_year)
  
```
