---
title: "pre_lab_01.Rmd"
author: "Derek Willis"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

## About this notebook

This notebook contains code and explanatory text that your should review and run as you read through chapters 13 and 15 of the course textbook, "Data Journalism with R and the Tidyverse". Answer questions and edit the document as directed.

Running this notebook will help you understand key data analysis methods and concepts that you will put into practice during this week's lab. When you come across a new command or verb in R, you should add it to your class reference notebook file along with what it's used for.

When you are finished running the code in this notebook, you will push changes to your course GitHub repo, and upload the link to ELMS as instructed.

You also should update your class reference notebook with code examples and what they mean. When you come across a new concept or R verb, put it in your reference notebook along with a sentence about what it does. For example:

```{r}
library(tidyverse)
```

The library function loads an R library such as the tidyverse, providing access to the verbs and functions that library contains. You have to load a library before you can use its functions.

## Instructor in class demonstration outline

-   Load this notebook, and provide overview of R, RStudio (text editor, console, environment window, file explorer), Rmarkdown data notebooks, the reference notebook.
-   Work through the code below, chapter by chapter.
    -   Chapter 13: explain the console and the concept of variables. Explain packages, the Tidyverse, how to install them.
    -   Chapter 15: explain loading libraries, the concept of functions, loading data, the importance of having questions to guide data analysis, and aggregation and summarization.

## Chapter 13

### Task 1: The console

**Task** In the console (look below), type 2+2 and hit enter. What does it return? Type your answer below. **Answer: 4**

### Task 2: Variables

**Task** Put the following code -- which stores the numeral 2 as a variable called number -- in the console and hit enter.

```{r}
number <- 2
```

Then type this code to multiply the variable called number by itself.

```{r}
number * number
```

What does it return? Type your answer below:

**Answer: 4**

### Task 3: Install packages

**Task** In the console, run the following code to install the Tidyverse:

```{r}
install.packages('tidyverse')
```

**Task** Add that

Then run the following code to install the rmarkdown library

```{r}
install.packages('rmarkdown')
```

### Task 4: Load libraries

**Task** Run the following code in the gray-colored codeblock below -- not in the console -- to load the tidyverse library. To run the code, click the little green play button (left facing arrow) at the top right of the codeblock. In Rmarkdown data notebooks, we write code inside of codeblocks, and explanatory text in the white area outside of it.

```{r}
# This is a comment inside of a codeblock. Comments doesn't run when we run the codeblock.
# Adding this command is a good way to start your class reference notebook.
library(tidyverse)
```

### Task 5: Load data

**Task** Load the UMD courses data by running the following codeblock. The data is contained in an .rds file, which is a binary format for R data.

```{r}
umd_courses <- read_rds("umd_courses.rds")
```

Once it loads, look at the environment window at right in RStudio for umd_courses It's next to a small blue circle with a white arrow in it. To the right of that, look at how many "obs" the dataset has ("obs" is short for observation, which is another way of saying rows or records). Look at how many variables it has (another way of saying columns or fields). Write the number of observations and variables below.

**Answer: 55632 rows, 7 columns**

### Task 6: Glimpse data

**Task** Run the following codeblock to use glimpse() to get a sense of the data, some of the values in each field/variable/column, and the data types of each field/variable/column.

```{r}
glimpse(umd_courses)
```

In the space below, write the data type for the "id" column and the "seats" column (it's in light gray, offset by \< \>). 
**Answer: "id" column is <chr>, "seats" column is <dbl>**

### Task 7: Examine data "head"

**Task** Run the following codeblock to use head() to examine the first six rows of data. Explore the data by clicking the black arrow on the right side of the output table, allowing you to examine additional columns.

```{r}
head(umd_courses)
```

In the space below, write the name of the business from the first row of the "name" column. 
**Answer: Directed Research**

### Task 8: Explore data with interactive dataframe

**Task** look at the environment window at right in RStudio for umd_courses. It's next to a small blue circle with a white arrow in it. Click on the umd_courses -- not the blue circle. It will open the dataframe as an interactive table. Scroll down to see additional rows. Scroll left to see additional columns. Click on a column name to sort by that column.

### Task 9: Group and count

**Task** Run the following codeblock:

```{r}
umd_courses %>%
  group_by(term) %>%
  summarise(
    count_classes = n()
  )
```

Describe what you think each line of this code is doing below, based on the result:

**Answer: The first line is selecting the dataset to look at, the second line is grouping the data by a specific condition, and the third and fourth lines are telling us the number of classes fit under each grouping condition, which is the term they're being offered in this case.**

### Task 10: Adding arrange to sort the data

**Task** Run the following codeblock:

```{r}
umd_courses %>%
  group_by(term) %>%
  summarise(
    count_classes = n()
  ) %>%
  arrange(desc(count_classes))
```

Compared to the previous codeblock, what has changed and how did that affect the result? 
**Answer: This code arranged the data in descending order by how many classes were offered during each term. The data was first summarized into the number of classes offered each term, then arranged into descending order.**

### Task 11: Grouping by multiple columns

**Task** Run the following codeblock.

```{r}
umd_courses %>%
  group_by(term, department) %>%
  summarise(
    count_classes = n()
  ) %>%
  arrange(term, department)
```

Compared to the previous codeblock, how did the output change and what in the code changed it?

**Answer: This codeblock added another condition to group data — the department that classes are in — which adjusted how the data was grouped. The output now shows the number of classes under each department and specific term.**

### Task 12: Other summary functions

**Task** Run the following codeblock:

```{r}
umd_courses %>%
  group_by(term) %>%
  summarise(
    count_classes = n(),
    total_seats = sum(seats)
  ) %>%
  arrange(desc(total_seats))
```

Compared to the previous codeblock, what has changed and how did that affect the result? 
**Answer: This took away department as a condition for grouping and instead added another metric to summarize the data: the total number of seats offered. The output shows one column for the total number of classes in a given term and another column for how many seats there are in a given term.**

### Task 13: different kinds of summarizations

**Task** Run the following codeblock to group the umd_courses dataframe by term. Then do the following summarizations for each term: count the number of courses, add up the total of all seats, and calculate the mean and median contribution number of seats. Then sort so the term with the most total seats is listed first.

```{r}
umd_courses %>%
  group_by(term) %>%
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats),
    median_seats = median(seats)
  ) %>%
  arrange(desc(total_seats))
```

Compared to the previous codeblock, what has changed and how did that affect the result? What do the top terms have in common? 
**Answer: In addition to the total number of classes and total number of seats for a given semester, the 7th and 8th lines of code add two more metrics, the mean number of seats per term and median number of seats per term, to summarize the data. There are now two more columns, one each for the mean and median number of seats in a given term. The mean number of seats for each term are much larger than the median number of seats.**

### Task 14: more aggregations

**Task** Run the following codeblock to group the umd_courses dataframe by the department. Then do the following summarizations for each group: count the number of courses, add up the total of all seats, calculate the mean and median number of seats and calculate the smallest and largest seat numbers. Then sort so the department with the largest total number of seats is listed first.

```{r}
umd_courses %>%
  group_by(department) %>%
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats),
    median_seats = median(seats),
    min_seats = min(seats),
    max_seats = max(seats)
  ) %>%
  arrange(desc(total_seats))
```

Compared to the previous codeblock, what has changed and how did that affect the result? How is the data sorted?

**Answer: Two more summarization metrics — the minimum number of seats available and maximum number of seats available — were added to the data output, and the data was also grouped by department rather than term. The data is sorted in descending order, starting from the department with the highest number of seats and decreasing from there.**

### Task 12: Questions

No code here; I want you to provide three questions that this data suggests to you. What would you like to ask it?

**Answer**

1.  How many departments have a max number of seats above 1000, and how many departments have a max number of seats below 1000?

2.  Which departments have the highest number of professors?

3.  Which professors teach the highest number of classes, and which department are they in?
